\boldmath Primordial Black Hole Formation in a Dust Bouncing Model

### 2024-12-03 main.tex
Removed font-color marking modifications.

{\color{red} We start by applying the aforementioned considerations to a flat,
	homogeneous, and isotropic universe containing a single perfect fluid characterized by
	its pressure $\bar{p}$ and energy density $\bar{\rho}$, along with the equation of state
	$\bar{p} = w\bar{\rho}$, with a constant $w$. Subsequently, we specialize in the dust
	case where $w \approx 0$. In practice, we consider a barotropic fluid with a constant
	equation of state, following the Schutz formalism as described in
	Ref.~\cite{fluidgeral}. In this section we will briefly review the quantization of the
	flat-dust bouncing model, focusing on the background dynamics.

@@

	In the above framework, the total Hamiltonian takes the form
	\begin{align}
		\label{total_hamiltonian_FLRW}
		{\cal H} & = \frac{N}{\bar{a}^{3w}}\dpar{\Pi_{T}- \frac{\kappa\Pi_q^{2} }{12}} ,
	\end{align}
	Canonical quantization is then performed by promoting classical variables to operators
	satisfying the canonical commutation relations. This process yields the following
	Wheeler-DeWitt equation for the wave-function of the universe  $\Psi(\bar{a}, T)$
	(see Ref.~\cite{nelson_peter_bouncing_original}):
	\begin{equation}
		\label{wdweq}
		i\hbar\frac{\partial}{\partial T}\Psi(q,T) -
		\frac{\kappa\hbar^2}{12}\frac{\partial^{2}}{\partial q^{2}}\Psi(q,T)  = 0 ,
	\end{equation}
	such that a specific operator factor ordering was chosen to preserve the symmetries of
	the classical system \cite{halliwell1990introductory}. Also, note that our last
	canonical transformation leads to the following variable
	\begin{equation}
		q = \frac{2\bar{a}^{\frac{3}{2}\dpar{1-w} } }{ 3\dpar{1-w } }.
	\end{equation}
	In the WdW equation above, we circumvent the problem of time by interpreting the
	parameter $T$ as an intrinsic time variable that is monotonically related to the
	classical cosmic time in the classical limit. This allows us to interpret the equation
	as a Schrödinger-like equation.
}

Equation~\eqref{wdweq} resembles a time-reversed free particle Schrödinger equation and,
with appropriate boundary conditions, its solutions are wave functions of the scale
factor $\bar{a}$. We turn to the De Broglie-Bohm
interpretation~\cite{mukhanov2005physical, nelson2000bohm}, such that assuming a Gaussian
wave-function $\Psi(q, T)$, the Bohmian trajectory solution translates to the scale
factor reads
\begin{equation}
	\label{bohm_scale_factor}
	\bar{a}(T) = \bar{a}_{B}\dcol{1 + \dpar{ \frac{T}{T_{ B}} }^2 }^{{\frac{1}{3}\frac{1}{\dpar{1-w}} } }  ,
\end{equation}
where $\bar{a}_{B}$ is an integration constant that represents the minimum scale factor
value and $T_B$ is a small arbitrary constant related to the time scale of the bounce.
	{\color{red} To simplify the analysis, we choose the lapse function $N = \bar{a}^{3w}$,
		but this choice is only possible after solving the Bohmian trajectory. The reason is
		that, initially, we cannot select $N$ as an operator function of the scale factor,
		However, once we have the Bohmian trajectory $\bar{a}(T)$, we can make this choice,
		using it purely for convenience in the analysis. In our model, $T$ extends from $T =
			-\infty$ in the far past to $T = 0$ at the bounce, where the scale factor $\bar{a}$
		reaches its minimum. } Remarkably, $\bar{a}(T) \neq 0$ for all $T$, which means that
this model is non-singular and represents an eternal
universe~\cite{nelson_peter_bouncing_original}. For a detailed derivation, see
\cite{nelson2021bouncing}.

{\color{red} It is important to clarify that in Ref.~\cite{nelson2021bouncing}, the
author selects a Gaussian wave function for the scale factor at the bounce. However,
this does not imply that this choice is the initial condition for our model. Generally,
a Gaussian wave packet of a one-dimensional free particle will develop a time-dependent
phase proportional to $q^2$. Thus, it is always possible to begin with a Gaussian wave
function at any time and define the bounce as the moment when this phase factor is zero.
For instance, in Ref.~\cite{Peter2016a}, the authors choose a Gaussian wave function at
any time and fix the phase using $H_0$. }

@@

Eliminating the time $t$ using \eqref{bohm_time} in \eqref{bohm_hubble}, we find
\begin{equation}
	\label{friedman_scale}
	\bar{H}^{2} = \frac{ 4 }{ 9t^{2}_{b}}\dpar
	{ \frac{ \bar{a}^{3}_{b} }{ \bar{a}^{3} } - \frac{\bar{a}_{b}^{6}}{ \bar{a}^{6} } }\, ,
\end{equation}
which is equivalent to the Friedmann equations
\begin{equation}\label{effective_friedmann}
	\bar{H}^{2} = \frac{\kappa}{ 3 }\bar\rho - \bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6},
\end{equation}
where $\bar\rho$ is the dust energy density. {\color{red} Also, from the derivative of
		Eq.~\eqref{friedman_scale},
		\begin{align}
			\dot{\bar{H}} = - \frac{\kappa}{2} \bar\rho + 3\bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}.
		\end{align}
	}

In the above equations, {\color{red} the overdot represents the time derivative and} there is an additional term
$-\bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}$ when compared to the usual Friedmann equation
for a classical universe. {\color{red} In our model, the scale factor dynamics resemble
		a typical Friedmann equation with a total energy density $\rho_{T} = \bar{\rho} +
			\rho_{q},$ where $\bar{\rho}$ is the dust energy density, and $\rho_{q} = -\Omega
			\rho_{c} \bar{a}^{-6}$ denotes an effective negative energy density with an equation of
		state $w_q = 1$ that accounts for quantum effects to avoid the primordial
		singularity~\cite{vitenti2012large}. It is important to note that $\rho_{q}$ is not a
		physical energy density but rather an effective one that emerges from the Bohmian trajectory. In
		the perturbative analysis, the energy density will always be associated only with
		$\bar{\rho}$. For large-scale factors $\bar{a} \gg \bar{a}_{B} $,
		Eq.~\eqref{effective_friedmann} and Eq.~\eqref{bohm_scale_factor} reduce to the
		classical dust dominated a flat FLRW universe. With this, we conclude the analysis of our
		background quantum bouncing model and move to its associated perturbations. In the next
		sections, we will consider the above quantum bouncing model with $w\approx 10^{-10}$ and
		$\bar{a}(t_0)/a_B = 10^{35}$ where $t_0$ is the time during the contracting phase where
		$H(t_0) = -H_0$ for $H_0 = 70 \, \text{km/s/Mpc}$. Moreover, when solving the equations
		numerically, we use the exact expressions in terms of $w$ without approximation.}

@@

	{\color{red} In the perturbative treatment, we continue to use the Schutz formalism
		and construct the Lagrangian as in Eq.~\eqref{lagrangian}, but now incorporating the
		perturbed metric. The perturbed Lagrangian and associated variables are defined as
		in Ref.~\cite{Vitenti2012}. Note that in the perturbed Lagrangian, the perturbations
		are always linked to the physical quantities and their corresponding background
		values. The effective energy density and pressure do not fluctuate; instead, the
		quantum effects from the background are incorporated through the use of the
		background Bohmian trajectories.}

@@

Using the background Einstein equations, we can relate the density contrast to the
Bardeen variables and the gauge invariant curvature perturbation $\zeta$ through the
following relations~\cite{Mukhanov1992, Vitenti2013}
\begin{align}
	\Psi                                                    & =\Phi,
	\\
	\label{deltarho}
	-\frac{2\bar{D}^2 \Phi}{3  \kappa (\bar{\rho}+\bar{p})} & =   \frac{{\tilde{\delta\rho}}} {3(\bar{\rho}+\bar{p})}
	,
	\\
	\label{vrelation}
	\zeta                                                   & \equiv \Psi  + \bar{H} \tilde{\mathcal{V}}
	\\
	\label{zeta2}
	\zeta                                                   & = \frac{3  \bar{a}^3}{N^2z^2 c_s^2 } \left[\frac{\partial}{\partial t}\left(\frac{\Phi }{3\bar{H}}\right) + \frac{\Phi}{3}\right]
	,                                                                                                                                                                                           \\
	\label{phiz}
	\bar{D}^{2} \Phi                                        & =  - \frac{z^2\bar{H}}{\bar{a}^3} \dot{\zeta} = -\frac{\bar{H}}{2\bar{a}^3} \Pi_\zeta
	.\end{align}
\textcolor{red}{In the above, $c^{2}_s=\dpar{ \frac{\partial\bar{\rho}}{\partial\bar{p}} }_{S}= w$ is the speed of sound, $\Pi_\zeta$ is the conjugated momenta related to $\zeta$ and
	\begin{align}
		\label{zdef}
		z^2=\frac{\kappa \bar{a}^3 (\bar{\rho} + \bar{p})}{2\bar{H}^2 c_{s}^2}
		.\end{align}}
For completeness, we also write the relation between the curvature perturbation and the
usual Mukhanov-Sasaki variable $v$, that is,
\begin{align}
	\label{msv}
	v & \equiv - \zeta z \sqrt{\frac{2}{\kappa}}
	.\end{align}
In the next section, we will establish a connection between PBH formation and density contrast. For now, it suffices to understand that the excess energy density associated with the perturbations leads to black hole formation, and we can measure such excess through the density contrast. Thus our definition in Eq.~\eqref{deltarhoinvariant} is extremely important and we will now analyze it.

@@

In this model, the background is characterized by a quantum contracting phase where an
effective quantum fluid dominates near the bounce. Consequently, we seek quantized
perturbations for consistency, although both theories can be viewed independently. In
particular, the use of quantized perturbations evolving on classical backgrounds has
been widely used since the results of Mukhanov and Chibisov \cite{mukhanov1981quantum}
and Hawking \cite{hawking1982quantum_fluctuations} to derive the power spectrum of
primordial perturbations, which are in turn used to describe the formation of structure
in our universe. {\color{red} In our model, we quantize both the background and the
		perturbations. Since we adopt the De Broglie-Bohm interpretation for the background,
		quantum effects are described in terms of the Bohmian scale factor, as shown in
		Eq.~\eqref{bohm_scale_factor}. This allows us to treat quantum perturbations evolving on
		this scale factor. To demonstrate the consistency of this approach,
		Ref.~\cite{Peter2005} derived the second-order Lagrangian for tensor perturbations
		without assuming classical equations of motion for the background. Furthermore, by using
		a factorized wave function for the perturbations and the background, the authors showed
		that these perturbations evolve conditioned on the Bohmian scale factor. In
		\cite{Vitenti2013}, the equivalent Lagrangian for scalar fluid perturbations was
		derived and is the one we use in this work.}

@@

Using these initial conditions for the modes, a well-defined vacuum state is obtained,
which is in turn used to construct the Fock Space by successive applications of the
creation operator $\hat{a}^{\dagger}_{k}$ and their linear combinations
\cite{mukhanov2007introduction}. Now that we have defined the appropriate initial
conditions for quantization, we may use them to solve the equations of motion in
Eq.~\eqref{eomosc} and obtain the mode functions, which we shall do numerically.
	{\color{red} Finally, it is important to note that the adiabatic vacuum prescription is
		consistent with the perturbative analysis of the model, as it leads to small
		perturbations as discussed in Ref.~\cite{vitenti2012large}.}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.6\textwidth]{deltac_jh_vs_tau.pdf}
	\caption{ Plot of the critical threshold vs time computed with the NumCosmo library for $w = 10^{-10}$ and $x_b = 10^{35}$. Each plot refers to a different scale $k$, while points on the same plot represents a different initial time for the collapse. The time interval for each mode starts at the Jeans scale, leading to the smallest threshold values, and ends for initial times when the perturbations reach the Hubble scale, which leads to higher threshold values. The time parameter in the x-axis is given by Eq.~\eqref{tauparameter}.}
	\label{deltacfig}
\end{figure}
\newpage
\subsection{$\bar{p} = w \bar{\rho}, \quad w \ll 1$}
\textcolor{red}{We want to analyze what would happen if the dark matter had such a small equation of state that could be thought as a presureless fluid.} In this case, Eq.~\eqref{deltacf} tells us that, for every scale $k$, we can always find a sufficient time in the past such that $x_{ini} \ll 1$, leading to the collapse of all density perturbations. In other words, if the universe is old enough, all perturbations will eventually collapse into a black hole until the universe reaches the bounce. If there was no bounce, the universe would completely collapse into black holes and our model would be highly unstable.

@@

\textcolor{red}{In this work, we have investigated the formation of primordial black
	holes within a Friedmann background, incorporating scalar perturbations. An alternative
	approach would be to explore PBH formation in a homogeneous yet anisotropic background.
	The presence of anisotropy could give rise to PBHs with significant angular momentum,
	contrasting with those formed in standard inflationary scenarios. A study of this
	possibility is left for future work}. The next phase of this research would be to apply
the same methodology to a universe populated with radiation and dark matter, which more
closely resembles the physical universe and the interaction between both fluids may
affect PBH production. If such a model proves reasonable, it would also be interesting
to improve the computation of the LTB model in App.~\ref{appc} for larger values of the
equation of state. Despite the Jeans scale being a good approximation, the introduction
of a non-pressureless fluid would also require the study of the pressure impact in our
characteristic scales. Furthermore, in a radiation-dominated quantum bouncing model, one
expects a greater blue tilt on the spectral index, which may affect the density contrast
modes for smaller scales. Said effect could lead to a larger formation of PBHs. This
analysis is left for future work.

>>>>

We start by applying the aforementioned considerations to a flat, homogeneous, and
isotropic universe containing a single perfect fluid characterized by its pressure
$\bar{p}$ and energy density $\bar{\rho}$, along with the equation of state $\bar{p} =
	w\bar{\rho}$, with a constant $w$. Subsequently, we specialize in the dust case where $w
	\approx 0$. In practice, we consider a barotropic fluid with a constant equation of
state, following the Schutz formalism as described in Ref.~\cite{fluidgeral}. In this
section we will briefly review the quantization of the flat-dust bouncing model,
focusing on the background dynamics.

@@

In the above framework, the total Hamiltonian takes the form
\begin{align}
	\label{total_hamiltonian_FLRW}
	{\cal H} & = \frac{N}{\bar{a}^{3w}}\dpar{\Pi_{T}- \frac{\kappa\Pi_q^{2} }{12}} ,
\end{align}
Canonical quantization is then performed by promoting classical variables to operators
satisfying the canonical commutation relations. This process yields the following
Wheeler-DeWitt equation for the wave-function of the universe  $\Psi(\bar{a}, T)$ (see
Ref.~\cite{nelson_peter_bouncing_original}):
\begin{equation}
	\label{wdweq}
	i\hbar\frac{\partial}{\partial T}\Psi(q,T) -
	\frac{\kappa\hbar^2}{12}\frac{\partial^{2}}{\partial q^{2}}\Psi(q,T)  = 0 ,
\end{equation}
such that a specific operator factor ordering was chosen to preserve the symmetries of
the classical system \cite{halliwell1990introductory}. Also, note that our last
canonical transformation leads to the following variable
\begin{equation}
	q = \frac{2\bar{a}^{\frac{3}{2}\dpar{1-w} } }{ 3\dpar{1-w } }.
\end{equation}
In the WdW equation above, we circumvent the problem of time by interpreting the
parameter $T$ as an intrinsic time variable that is monotonically related to the
classical cosmic time in the classical limit. This allows us to interpret the equation
as a Schrödinger-like equation.

Equation~\eqref{wdweq} resembles a time-reversed free particle Schrödinger equation and,
with appropriate boundary conditions, its solutions are wave functions of the scale
factor $\bar{a}$. We turn to the De Broglie-Bohm
interpretation~\cite{mukhanov2005physical, nelson2000bohm}, such that assuming a Gaussian
wave-function $\Psi(q, T)$, the Bohmian trajectory solution translates to the scale
factor reads
\begin{equation}
	\label{bohm_scale_factor}
	\bar{a}(T) = \bar{a}_{B}\dcol{1 + \dpar{ \frac{T}{T_{ B}} }^2 }^{{\frac{1}{3}\frac{1}{\dpar{1-w}} } }  ,
\end{equation}
where $\bar{a}_{B}$ is an integration constant that represents the minimum scale factor
value and $T_B$ is a small arbitrary constant related to the time scale of the bounce.
To simplify the analysis, we choose the lapse function $N = \bar{a}^{3w}$, but this
choice is only possible after solving the Bohmian trajectory. The reason is that,
initially, we cannot select $N$ as an operator function of the scale factor, However,
once we have the Bohmian trajectory $\bar{a}(T)$, we can make this choice, using it
purely for convenience in the analysis. In our model, $T$ extends from $T = -\infty$ in
the far past to $T = 0$ at the bounce, where the scale factor $\bar{a}$ reaches its
minimum. Remarkably, $\bar{a}(T) \neq 0$ for all $T$, which means that this model is
non-singular and represents an eternal universe~\cite{nelson_peter_bouncing_original}.
For a detailed derivation, see~\cite{nelson2021bouncing}.

It is important to clarify that in Ref.~\cite{nelson2021bouncing}, the author selects a
Gaussian wave function for the scale factor at the bounce. However, this does not imply
that this choice is the initial condition for our model. Generally, a Gaussian wave
packet of a one-dimensional free particle will develop a time-dependent phase
proportional to $q^2$. Thus, it is always possible to begin with a Gaussian wave
function at any time and define the bounce as the moment when this phase factor is zero.
For instance, in Ref.~\cite{Peter2016a}, the authors choose a Gaussian wave function at
any time and fix the phase using $H_0$.

@@

Eliminating the time $t$ using \eqref{bohm_time} in \eqref{bohm_hubble}, we find
\begin{equation}
	\label{friedman_scale}
	\bar{H}^{2} = \frac{ 4 }{ 9t^{2}_{b}}\dpar
	{ \frac{ \bar{a}^{3}_{b} }{ \bar{a}^{3} } - \frac{\bar{a}_{b}^{6}}{ \bar{a}^{6} } }\, ,
\end{equation}
which is equivalent to the Friedmann equations
\begin{equation}\label{effective_friedmann}
	\bar{H}^{2} = \frac{\kappa}{ 3 }\bar\rho - \bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6},
\end{equation}
where $\bar\rho$ is the dust energy density. Also, from the derivative of
Eq.~\eqref{friedman_scale},
\begin{align}
	\dot{\bar{H}} = - \frac{\kappa}{2} \bar\rho + 3\bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}.
\end{align}

In the above equations, the overdot represents the time derivative and there is an additional term
$-\bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}$ when compared to the usual Friedmann equation
for a classical universe. In our model, the scale factor dynamics resemble
a typical Friedmann equation with a total energy density $\rho_{T} = \bar{\rho} +
	\rho_{q},$ where $\bar{\rho}$ is the dust energy density, and $\rho_{q} = -\Omega
	\rho_{c} \bar{a}^{-6}$ denotes an effective negative energy density with an equation of
state $w_q = 1$ that accounts for quantum effects to avoid the primordial
singularity~\cite{vitenti2012large}. It is important to note that $\rho_{q}$ is not a
physical energy density but rather an effective one that emerges from the Bohmian
trajectory. In the perturbative analysis, the energy density will always be associated
only with $\bar{\rho}$. For large-scale factors $\bar{a} \gg \bar{a}_{B} $,
Eq.~\eqref{effective_friedmann} and Eq.~\eqref{bohm_scale_factor} reduce to the
classical dust dominated a flat FLRW universe. With this, we conclude the analysis of
our background quantum bouncing model and move to its associated perturbations. In the
next sections, we will consider the above quantum bouncing model with $w\approx
	10^{-10}$ and $\bar{a}(t_0)/a_B = 10^{35}$ where $t_0$ is the time during the
contracting phase where $H(t_0) = -H_0$ for $H_0 = 70 \, \text{km/s/Mpc}$. Moreover,
when solving the equations numerically, we use the exact expressions in terms of $w$
without approximation.

@@

In the perturbative treatment, we continue to use the Schutz formalism and construct the
Lagrangian as in Eq.~\eqref{lagrangian}, but now incorporating the perturbed metric. The
perturbed Lagrangian and associated variables are defined as in Ref.~\cite{Vitenti2012}.
Note that in the perturbed Lagrangian, the perturbations are always linked to the
physical quantities and their corresponding background values. The effective energy
density and pressure do not fluctuate; instead, the quantum effects from the background
are incorporated through the use of the background Bohmian trajectories.

@@

Using the background Einstein equations, we can relate the density contrast to the
Bardeen variables and the gauge invariant curvature perturbation $\zeta$ through the
following relations~\cite{Mukhanov1992, Vitenti2013}
\begin{align}
	\Psi                                                    & =\Phi,
	\\
	\label{deltarho}
	-\frac{2\bar{D}^2 \Phi}{3  \kappa (\bar{\rho}+\bar{p})} & =   \frac{{\tilde{\delta\rho}}} {3(\bar{\rho}+\bar{p})}
	,
	\\
	\label{vrelation}
	\zeta                                                   & \equiv \Psi  + \bar{H} \tilde{\mathcal{V}}
	\\
	\label{zeta2}
	\zeta                                                   & = \frac{3  \bar{a}^3}{N^2z^2 c_s^2 } \left[\frac{\partial}{\partial t}\left(\frac{\Phi }{3\bar{H}}\right) + \frac{\Phi}{3}\right]
	,                                                                                                                                                                                           \\
	\label{phiz}
	\bar{D}^{2} \Phi                                        & =  - \frac{z^2\bar{H}}{\bar{a}^3} \dot{\zeta} = -\frac{\bar{H}}{2\bar{a}^3} \Pi_\zeta
	.\end{align}
In the above, $c^{2}_s=\dpar{ \frac{\partial\bar{\rho}}{\partial\bar{p}} }_{S}= w$ is
the speed of sound, $\Pi_\zeta$ is the conjugated momenta related to $\zeta$ and
\begin{align}
	\label{zdef}
	z^2=\frac{\kappa \bar{a}^3 (\bar{\rho} + \bar{p})}{2\bar{H}^2 c_{s}^2}.
\end{align}
For completeness, we also write the relation between the curvature perturbation and the
usual Mukhanov-Sasaki variable $v$, that is,
\begin{align}
	\label{msv}
	v & \equiv - \zeta z \sqrt{\frac{2}{\kappa}}
	.\end{align}
In the next section, we will establish a connection between PBH formation and density contrast. For now, it suffices to understand that the excess energy density associated with the perturbations leads to black hole formation, and we can measure such excess through the density contrast. Thus our definition in Eq.~\eqref{deltarhoinvariant} is extremely important and we will now analyze it.

@@

In this model, the background is characterized by a quantum contracting phase where an
effective quantum fluid dominates near the bounce. Consequently, we seek quantized
perturbations for consistency, although both theories can be viewed independently. In
particular, the use of quantized perturbations evolving on classical backgrounds has
been widely used since the results of Mukhanov and Chibisov \cite{mukhanov1981quantum}
and Hawking \cite{hawking1982quantum_fluctuations} to derive the power spectrum of
primordial perturbations, which are in turn used to describe the formation of structure
in our universe. In our model, we quantize both the background and the
perturbations. Since we adopt the De Broglie-Bohm interpretation for the background,
quantum effects are described in terms of the Bohmian scale factor, as shown in
Eq.~\eqref{bohm_scale_factor}. This allows us to treat quantum perturbations evolving on
this scale factor. To demonstrate the consistency of this approach,
Ref.~\cite{Peter2005} derived the second-order Lagrangian for tensor perturbations
without assuming classical equations of motion for the background. Furthermore, by using
a factorized wave function for the perturbations and the background, the authors showed
that these perturbations evolve conditioned on the Bohmian scale factor. In
\cite{Vitenti2013}, the equivalent Lagrangian for scalar fluid perturbations was
derived and is the one we use in this work.

@@

Using these initial conditions for the modes, a well-defined vacuum state is obtained,
which is in turn used to construct the Fock Space by successive applications of the
creation operator $\hat{a}^{\dagger}_{k}$ and their linear combinations
\cite{mukhanov2007introduction}. Now that we have defined the appropriate initial
conditions for quantization, we may use them to solve the equations of motion in
Eq.~\eqref{eomosc} and obtain the mode functions, which we shall do numerically.
Finally, it is important to note that the adiabatic vacuum prescription is consistent
with the perturbative analysis of the model, as it leads to small perturbations as
discussed in Ref.~\cite{vitenti2012large}.

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.6\textwidth]{deltac_jh_vs_tau.pdf}
	\caption{ Plot of the critical threshold vs time computed with the NumCosmo library for $w = 10^{-10}$ and $x_b = 10^{35}$. Each plot refers to a different scale $k$, while points on the same plot represents a different initial time for the collapse. The time interval for each mode starts at the Jeans scale, leading to the smallest threshold values, and ends for initial times when the perturbations reach the Hubble scale, which leads to higher threshold values. The time parameter in the x-axis is given by Eq.~\eqref{tauparameter}.}
	\label{deltacfig}
\end{figure}
\newpage
\subsection{$\bar{p} = w \bar{\rho}, \quad w \ll 1$}

We want to analyze what would happen if the dark matter had such a small equation of
state that could be thought as a presureless fluid. In this case, Eq.~\eqref{deltacf}
tells us that, for every scale $k$, we can always find a sufficient time in the past
such that $x_{ini} \ll 1$, leading to the collapse of all density perturbations. In
other words, if the universe is old enough, all perturbations will eventually collapse
into a black hole until the universe reaches the bounce. If there was no bounce, the
universe would completely collapse into black holes and our model would be highly
unstable.

@@

In this work, we have investigated the formation of primordial black holes within a
Friedmann background, incorporating scalar perturbations. An alternative approach would
be to explore PBH formation in a homogeneous yet anisotropic background. The presence of
anisotropy could give rise to PBHs with significant angular momentum, contrasting with
those formed in standard inflationary scenarios. A study of this possibility is left for
future work. The next phase of this research would be to apply the same methodology to a
universe populated with radiation and dark matter, which more closely resembles the
physical universe and the interaction between both fluids may affect PBH production. If
such a model proves reasonable, it would also be interesting to improve the computation
of the LTB model in App.~\ref{appc} for larger values of the equation of state. Despite
the Jeans scale being a good approximation, the introduction of a non-pressureless fluid
would also require the study of the pressure impact in our characteristic scales.
Furthermore, in a radiation-dominated quantum bouncing model, one expects a greater blue
tilt on the spectral index, which may affect the density contrast modes for smaller
scales. Said effect could lead to a larger formation of PBHs. This analysis is left for
future work.
